D3064R0: Bringing the Laws of Physics, Computer-System Constraints, and Compilers to Bear on the OOTA Problem

### 2024-01-29 ootavacuous.tex
FYI, possible progress on language-level OOTA problem
On Sun, Jan 28, 2024 at 04:37:33PM -0800, Paul E. McKenney wrote:
> > Do I need the c2latex.sh script to build this?  Whatever it is, I don't
> > have it.
>
> You do!  One-line awk script, but attached.
>
> I also moved it into the git archive and pushed it out.
>
> > Also, the \includegraphics LaTeX command doesn't seem to be able to work
> > with .svg files.  I get "File not found" errors.
>
> Ah, apologies, you need to build with "sh ootavacuous.sh".  Very crude,
> I admit.  I created and pushed a README file saying this.

All working no

OOTA occurs when a group of threads load from each others' stores to
each of a group of objects, but where each thread's store depends on
the value returned by that thread's load.
A given out-of-thin-air value passes around the resulting \emph{OOTA cycle}.

@@

However, it is important to distinguish OOTA cycles from the OOTA-like
behavior that can be caused by simple reordering.
An example of simple reordering is shown in
Listing~\ref{lst:Simple Reordering}~\cite{PaulEMcKenney2020RelaxedGuideRelaxed}.
Both the C++ compiler and the CPU are within their rights to reorder
lines~12 and~13, which can result in all of \co{x}, \co{y}, \co{r1},
and \co{r2} having the value 42 as follows:

@@

Current C++ implementations can and do exhibit this simple-reordering
behavior.

@@

In contrast,
Listing~\ref{lst:Simple OOTA}~\cite{PaulEMcKenney2020RelaxedGuideRelaxed}
shows a simple example where an OOTA cycle might result in all of \co{x},
\co{y}, \co{r1}, and \co{r2} have final values of 42, despite the fact
that there is nothing in the initial values or the executable code that
would support such an outcome:

\begin{enumerate}
\item   Line~6 loads from \co{x} into \co{r1}, somehow obtaining the
	value 42, never mind the \co{x}'s initial value of zero.
\item   Line~7 stores \co{r1}, and thus 42, to \co{y}.
\item   Line~12 loads 42 from \co{y} to \co{r2}.
\item   Line~13 stores 42 to \co{x}, justifying the value loaded by
	line~6 after the fact.
\end{enumerate}

Fortunately, the C++ standard prohibits such OOTA cycles in 33.5.4p8
(\co{[atomics.order]})~\cite{ThomasKoeppe2023N4950}:

@@

The standard's prohibition of OOTA is of course important, but those of us
writing code in the real world must rely on actual C++ implementations.
And in these implementations, this prohibition is in fact enforced by TSO
ordering in strongly ordered systems and by data dependency ordering in
weakly ordered systems.
Real-world implementations cannot store something until after they load
it, so there is a \emph{semantic dependency} from line~6 to line~7 and
another from line~12 to line~13 in
Listing~\ref{lst:Simple OOTA}.
These semantic dependencies cause compilers to emit instructions such
that the hardware enforcement applies.\footnote{
	The need to prohibit simple OOTA is one reason why compiler-based
	value speculation optimizations require checks on such
	speculation, and these checks must be based on actual values
	loaded.}

Roughly speaking, \emph{there is semantic dependency from a given load
to a given store when, all other things being equal a change in the
value loaded can change the value stored.}

This paper will follow P2055R0~\cite{PaulEMcKenney2020RelaxedGuideRelaxed}
in using the term \emph{full C++} to denote the standard including the
above prohibition, and the term \emph{strict C++}\footnote{
	@@@ Future bikeshedding of name.}
% @@@ Alternatives thus far: OOTA-prone C++, OOTA-enabled C++,
% @@@ relaxed C++, OOTA C++, OOTA-agnostic C++.
to denote a hypothetical standard that excludes this prohibition, but
is otherwise identical to full C++.
Unqualified \emph{C++} means full C++.

The next section looks at how prior work has refined this prohibition.

@@

Goldblatt looked at interactions between OOTA cycles and
undefined behavior (UB)~\cite{DavidGoldblatt2019NoElegantOOTAfix}, and
Appendix~\ref{app:Aside on Undefined Behavior}
analyzes his examples and notes ways to separate UB and OOTA-cycle
concerns.
The remainder of this document will focus on examples lacking UB.

All this work focused on either identifying OOTA or on how C++
implementations could avoid it.
None of this work applied real-world hardware ordering to the problem
of avoiding OOTA cycles.
Which might explain why no known real-world C++ implementation results
in OOTA executions.

This paper therefore digs more deeply into OOTA cycles in the light
of these real-world constraints.

@@

This paper will use the \co{herd7}\footnote{
	Available at \path{https://github.com/herd/herdtools7}.}
tool to analyze fragments of C++ code.
This tool carries out the moral equivalent of full state-space searches
of concurrent code fragments.
In some cases its output will include executions with OOTA cycles,
on occasion reporting undefined values for the variables involved in
the cycle.

@@

Listing~\ref{lst:OOTA Cycle}
shows a code fragment that in strict C++ would have an OOTA cycle (but
that is of course prohibited in full C++).
This section describes this fragment, thus giving an overview of
the \co{herd7} tool.

The first line identifies it as a C-language litmus test and gives it
a name.
Lines~2-5 initialize variables, in this case setting the initial
values of the global shared variables \co{x} and \co{y} to zero.
Lines~7-13 define the first thread, \co{P0()}, and lines~15-21
define the second thread, \co{P1()}.\footnote{
	The ``P'' stands for ``process'', which is \co{herd7}'s name
	for ``thread''.}
The arguments to both \co{P0()} and \co{P1()} specify which of
the global shared variables each thread may access, in this case, \co{x}
and \co{y}.
The body of each thread specifies a limited subset of C++ code,
which also happens to be a somewhat less limited subset of C code.

Line~23 gives an example of a \co{locations} clause, which causes
\co{herd7} to dump out the final value of \co{x} and \co{y}.

Finally, line~24 specifies an \co{exists} clause, which gives a
condition to check for the final value of the specified local
variables.
The \co{0:} prefix specifies a variable local to \co{P0()} and
the \co{1:} prefix specifies a variable local to \co{P1()}.
The \co{/\\} specifies a boolean AND, and the \co{=} specifies
an equality comparison.
If a variable appears in an \co{exists} clause, then the final value of
that variable constitutes observable behavior.

@@

Listing~\ref{lst:OOTA Cycle, herd7 Output} shows the corresponding output
of the \co{herd7} tool.
Lines~3 and~4 show the possible states, with line~4 showing the
counterintuitive outcome where both threads load the value 42.
Normally these states would include only those variables mentioned
in the \co{exists} clause, but because of the \co{locations} clause
on line~23, the values of \co{x} and \co{y} is also included, which
can be helpful when debugging.
% @@@ As we will see, this outcome is not possible in real-world
% @@@ implementations due to physical constraints of which \co{herd7}
% @@@ is (by design) unaware.
Line~9 contains \co{Sometimes} (as opposed to \co{Never} or \co{Always}),
indicating that some executions satisfy the \co{exists} clause and
others do not.

The earlier examples shown in
Listings~\ref{lst:Simple Reordering}
and~\ref{lst:Simple OOTA}
can be recast as similar \co{herd7} litmus tests, which areas shown in
Listing~\ref{lst:Simple Reordering as Litmus Test}
on page~\pageref{lst:Simple Reordering as Litmus Test}
and Listing~\ref{lst:Causality Test Case 4}
on page~\pageref{lst:Causality Test Case 4}, respectively.
Other OOTA-related litmus tests may be found in
Appendix~\ref{app:OOTA-Related Litmus Tests}.

\section{Semantic Dependencies, Executions, and OOTA}
\label{sec:Semantic Dependencies; Executions; and OOTA}

@@

However, semantic dependencies are only one type of link in an OOTA cycle.
The other link extends from a given store to the load that returns the
value stored.
It is tempting to make the intuitive argument that this store must precede
this load in global time, and then to combine this with the notion that
any real C++ implementation must consume global time when computing a
semantic dependency.
This combination suggests that such implementations cannot exhibit
OOTA cycles.
This combination has been formalized by defining an OOTA cycle as a cycle
in sdep $\cup$ rf, where sdep is the set of semantic dependencies within
each thread and rf is the set of store-to-load links, whether intra-thread
(rfi) or between threads (rfe).\footnote{
	See Appendix~\ref{app:Inter-Process Communications} for definitions
	and properies of rf, rfe, and rfi.}

This is a fine definition, and is consistent with the words in the C++
standard, but has an intra-thread problem exemplified by the following
code:

@@

This is an elaboration of \co{thread2()} from
Listing~\ref{lst:Simple OOTA}
that adds \co{z} along with lines~2 and~3.
The problem is that the C++ implementation is within its rights to
note that line~3 might well execute immediately after line~2, giving
threads no chance to modify \co{y}.
Such an implementation might therefore behave as if the source code
had instead been as follows:

@@

Here line~3 has been optimized away in favor of line~4 storing the value
in \co{r2} that was loaded from \co{y} by line~2.
And given that the load from \co{z} no longer exists, it cannot possibly
act as a temporal constraint.

In order to avoid these rfi links, we substitute rfe
for rf, thus defining an OOTA cycle as a cycle in sdep $\cup$ rfe (or
\co{acyclic(sdep ; rfe)} in \co{herd7} notation), where rfe is the set
of store-to-load links, but excluding intra-thread rfi links, which are
instead considered to be part of sdep.
Although this does shunt additional complexity onto the term
``semantic dependency'', it also enables us to cleanly separate
the inter-thread and intra-thread portions of any given OOTA cycle.
% \footnote{
%	Thanks to Alan Stern for providing a litmus test that demonstrated
%	that we had unwittingly (but productively!) shifted our definition
%	to sdep $\cup$ rfe.}

@@

Some discussions of semantic dependencies assume that semantic
dependencies are a strict function of the source code.
Although there are some ways of making this work, many uses of
semantic dependencies require them to be a function of particular
executions.
Consider for example:

@@

As long as \co{z} is zero, and all else being equal, changes in the
value of \co{y} will not cause changes to the value of \co{x}.
In that case, the implementation is within its rights to discard this
statement and act as if later accesses to \co{x} were instead the
constant zero, at least up to the next assignment to \co{x}.

@@

Suppose that both \co{y} and \co{z} are zero in the previous example.
Then changes to either but not both of \co{y} or \co{z} (again, all
else being equal) will not cause changes to the value of \co{x}.
In other words, there is no semantic dependency from either \co{y}
or \co{z} to \co{x} unless both \co{y} and \co{z} change.
Alternatively, when both \co{y} and \co{z} are zero, there is no semantic
dependency from either \co{y} or \co{z} to \co{x}, but rather a semantic
dependency from the combination of \co{y} \emph{and} \co{z} to \co{x}.
This means that precise definitions of sdep cannot always rely on
single-variable perturbations.
See Appendix~\ref{app:Non-Trivial Semantic Dependencies}
for examples and additional discussion.

Roughly speaking and all other things being equal, \emph{there is semantic
dependency from a given set of loads to a given store when a change in
the values loaded can change the value stored.}

This definition allows the definition of an OOTA cycle to be expanded
from \co{acyclic(sdep ; rfe)} to:

\begin{quote}
	There are stores
	$W_i (i = 0, ..., m)$,
	where $W_i$ semantically depends on loads
	$R_{i,j} (j = 0, ..., n_i)$,
	such that each $R_{i,j}$ reads from one of the $W_k$
	stores in a different thread.
\end{quote}

This definition makes OOTA more complicated than a simple cycle but
we will continue to refer to it as an OOTA cycle out of habit.
Note that this new definition includes and generalizes our previous
\co{acyclic(sdep ; rfe)} definition.

\subsubsection{Semantic Dependencies Affected by Optimizations}
\label{sec:Semantic Dependencies Affected by Optimizations}

@@

There are semantic dependencies from \co{y} to \co{x} and from \co{z}
to \co{x}.
However, if the implementation knows that the value of \co{y} is
always the negative of \co{z} there is no semantic dependency and the
implementation can again act as if later accesses to \co{x} were instead
the constant zero, again, at least until the next assignment to \co{x}.
Please note that this leaves aside the question of how the compiler
would know this if \co{y} and \co{z} are atomic objects that cannot be
updated pairwise atomically and are subject to change at any time by
other threads (a point we will return to).

@@

Consider the following simple \co{if} statement:

@@

Here there is a semantic dependency from \co{x}, but in some executions
it extends to \co{y} and in others to \co{z}.
This is an example of the load affecting not the value or address of
a given store, but rather whether or not that store is executed at all.

Which raises the question:
What exactly is an execution?

>>>>

In broad terms,
OOTA occurs when a group of threads load from each others' stores
and each thread's store depends on the value returned by that thread's load.
The collection of loads and stores forms an \emph{OOTA cycle}.
In the most extreme cases a nonsensical value can pop up ``out of thin air'',
but this is not necessary.

@@

It is important to distinguish true OOTA cycles from OOTA-like
behavior caused by simple reordering.
An example of simple reordering is shown in
Listing~\ref{lst:Simple Reordering}~\cite{PaulEMcKenney2020RelaxedGuideRelaxed}.
Both the C++ compiler and the CPU are within their rights to reorder
lines~12 and~13, which can result in all of \co{x}, \co{y}, \co{r1},
and \co{r2} having the value 42 as follows:

@@

\noindent
Current C++ implementations can and do exhibit this reordering behavior.

@@

In contrast,
Listing~\ref{lst:Simple OOTA}~\cite{PaulEMcKenney2020RelaxedGuideRelaxed}
shows a simple example where an OOTA cycle might result in all of \co{x},
\co{y}, \co{r1}, and \co{r2} having final values of 42, despite the fact
that there is nothing in the initial values or the executable code to
support such an outcome:

\begin{enumerate}
\item   Line~6 loads from \co{x} into \co{r1}, claiming to read
	the value of line~13's store rather than \co{x}'s initial value
	and somehow obtaining 42.
\item   Line~7 stores \co{r1}, and thus 42, to \co{y}.
\item   Line~12 loads 42 from \co{y} to \co{r2}.
\item   Line~13 stores 42 to \co{x}, justifying the value loaded by line~6.
\end{enumerate}

Because nothing else in the C++ memory model rules out such OOTA cycles,
the C++ standard explicitly prohibits them in 33.5.4p8
(\co{[atomics.order]})~\cite{ThomasKoeppe2023N4950}:

@@

\noindent
The standard's prohibition of OOTA is of course important, but those of us
writing code in the real world must rely on actual C++ implementations.
And in these implementations, this prohibition is in fact enforced by TSO
ordering in strongly ordered systems and by data dependency ordering in
weakly ordered systems.\footnote{
	The need to prohibit simple OOTA is one reason why compiler-based
	value speculation optimizations require checks on such
	speculation, and these checks must be based on actual values
	loaded.}

In Listing~\ref{lst:Simple OOTA}
there is a \emph{semantic dependency} from line~6 to line~7 and
another from line~12 to line~13.
(Roughly speaking, there is a semantic dependency from a given load
to a given store when \emph{all other things being equal, a change in the
value loaded can change the value stored or prevent the store from
occurring at all.}
Here the dependencies are trivial, because the values stored simply
\emph{are} the values that were loaded.)
Since real-world CPUs cannot store something
until they have determined its value,\footnote{
	Another way of saying this is that real-world CPUs do not
	speculate stores.}
the stores in lines~7 and~13 cannot take place until the CPU
knows what values are loaded by lines~6 and~12, respectively.
Thus the hardware orders these stores after their corresponding loads,
and this ordering prevents the OOTA result.

This paper will follow P2055R0~\cite{PaulEMcKenney2020RelaxedGuideRelaxed}
in using the term \emph{full C++} to denote the standard including the
prohibition above, and the term \emph{strict C++}\footnote{
	@@@ Future bikeshedding of name.}
% @@@ Alternatives thus far: OOTA-prone C++, OOTA-enabled C++,
% @@@ relaxed C++, OOTA C++, OOTA-agnostic C++.
to denote a hypothetical standard that excludes this prohibition, but
is otherwise identical to full C++.
Unqualified \emph{C++} means full C++.

The next section looks at how prior work has refined these issues.

@@

Goldblatt looked at interactions between OOTA cycles and
undefined behavior (UB)~\cite{DavidGoldblatt2019NoElegantOOTAfix}, and
Appendix~\ref{app:Aside on Undefined Behavior}
analyzes his examples and notes ways to separate UB and OOTA-cycle
concerns.
The remainder of this document will concentrate on examples lacking UB.

All this work focused on either identifying OOTA or seeing how C++
implementations could avoid it.
None applied real-world hardware ordering constraints to the problem
of avoiding OOTA cycles,
yet doing so might help explain why no known real-world C++ implementation
results in OOTA behavior.
We therefore dig more deeply into OOTA cycles in the light
of these real-world constraints.

@@

This paper will use the \co{herd7}\footnote{
	Available at \path{https://github.com/herd/herdtools7}.}
tool to analyze fragments of C++ code.
This tool carries out the moral equivalent of full state-space searches
of concurrent code fragments.
In some cases its output will include executions with OOTA cycles,
on occasion reporting undefined values for the variables involved in
the cycle.\footnote{
	This happens only some of the time because of idiosyncrasies
	in the algorithm used by \co{herd7}'s self-consistency solver.}

@@

Listing~\ref{lst:OOTA Cycle}
shows a code fragment that under strict C++ has an OOTA cycle (although
the cycle is of course prohibited in full C++).
This section describes the fragment, thereby giving an overview of
the \co{herd7} tool.

The first line identifies it as a C-language litmus test and gives it
a name.
Lines~2--5 initialize variables, in this case setting the initial
values of the global shared variables \co{x} and \co{y} to zero.
Lines~7--13 define the first thread, \co{P0()}, and lines~15--21
define the second thread, \co{P1()}.\footnote{
	The ``P'' stands for ``process'', which is \co{herd7}'s name
	for ``thread''.}
The arguments to both \co{P0()} and \co{P1()} specify which of
the global shared variables each thread may access, in this case, \co{x}
and \co{y}.
The body of each thread contains a limited subset of C++ code
(written in a slightly stilted manner to keep the load operations separate
from the rest).

Line~23 has a \co{locations} clause, which causes
\co{herd7} to dump out the final values of \co{x} and \co{y}.
Finally, line~24 specifies an \co{exists} clause, which gives a
condition to check for the final values of the specified variables.
The \co{0:} prefix denotes a variable local to \co{P0()} and
the \co{1:} prefix denotes a variable local to \co{P1()}.
The \co{/\\} is a boolean AND, and the \co{=} signs are equality comparisons.
If a variable appears in the \co{exists} clause then the final value of
that variable constitutes observable behavior.

@@

Listing~\ref{lst:OOTA Cycle, herd7 Output} shows the corresponding output
of the \co{herd7} tool.
Lines~3 and~4 show the possible states, with line~4 showing the
counterintuitive outcome where both threads load the value 42.
Normally these lines would include only those variables mentioned
in the \co{exists} clause, but because of the \co{locations} clause
the values of \co{x} and \co{y} are also listed, which
can be helpful for debugging.
% @@@ As we will see, this outcome is not possible in real-world
% @@@ implementations due to physical constraints of which \co{herd7}
% @@@ is (by design) unaware.
Line~9 contains \co{Sometimes} (as opposed to \co{Never} or \co{Always}),
indicating that some executions satisfy the \co{exists} clause and
others do not.

The earlier examples shown in
Listings~\ref{lst:Simple Reordering}
and~\ref{lst:Simple OOTA}
can be recast as similar \co{herd7} litmus tests, as shown in
Listing~\ref{lst:Simple Reordering as Litmus Test}
on page~\pageref{lst:Simple Reordering as Litmus Test}
and Listing~\ref{lst:Causality Test Case 4}
on page~\pageref{lst:Causality Test Case 4}, respectively.
Other OOTA-related litmus tests may be found in
Appendix~\ref{app:OOTA-Related Litmus Tests}.

\section{OOTA, Semantic Dependencies, and Executions}
\label{sec:OOTA; Semantic Dependencies; and Executions}

@@

Semantic dependencies are only one type of link in an OOTA cycle.
The other type extends from a given store to the load that returns the
value stored.
It is tempting to make the intuitive argument that the store must precede
the load in global time, and then to combine this with the notion that
any real C++ implementation must consume global time when computing a
semantic dependency.
This combination suggests that OOTA cycles cannot occur.
The idea has been formalized by defining an OOTA cycle as a cycle
in sdep $\cup$ rf [@@@ Reference needed],
where sdep is the set of semantic dependencies within
each thread and rf is the set of store-to-load ``reads-from'' links,
whether intra-thread
(rfi) or between threads (rfe).\footnote{
	See Appendix~\ref{app:Inter-Process Communications} for definitions
	and properies of rf, rfe, and rfi.}
This is a fine definition and is consistent with the words in the C++
standard, but it has an intra-thread problem exemplified by the following
code:

@@

\noindent
This is an elaboration of \co{thread2()} from
Listing~\ref{lst:Simple OOTA}
that adds \co{z} along with lines~3 and~4.
The problem is that a C++ implementation is within its rights to
note that line~3 might well execute immediately after line~2, giving
other threads no chance to modify \co{y} in between.
Such an implementation might therefore behave as if the source code
had instead been as follows:

@@

\noindent
Here line~3 has been optimized away in favor of line~4 storing the same value
in \co{r2} that was stored to \co{y} by line~2.
And given that the load from \co{y} no longer exists, it cannot possibly
act as a temporal constraint.

In order to avoid these rfi links we will substitute rfe
for rf, defining an OOTA cycle---for now---as a cycle in sdep $\cup$ rfe,
where rfe is the set of store-to-load links excluding intra-thread rfi links.
Any rfi links in a cycle can instead be interpreted as part of sdep.
Although this does shunt additional complexity onto the term
``semantic dependency'', it also enables us to separate cleanly
the inter-thread and intra-thread portions of any given OOTA cycle.
% \footnote{
%	Thanks to Alan Stern for providing a litmus test that demonstrated
%	that we had unwittingly (but productively!) shifted our definition
%	to sdep $\cup$ rfe.}

@@

Some discussions of semantic dependencies assume that they
are strictly functions of the source code.
Although there are ways of making this work, many instances of
semantic dependency must be considered functions of particular executions.
Consider for example:

@@

\noindent
As long as \co{z} is zero, and all else being equal, changes in the
value of \co{y} will not cause a change in the value stored to \co{x}.
As a result, the semantic dependency from \co{y} to \co{x} exists only
in executions where \co{z} is nonzero,
which shows it is a property of the execution, not just of the source code.

@@

Suppose that in some execution of the previous example,
both \co{y} and \co{z} are zero.
Then changes to either \co{y} or \co{z} (again, all
else being equal) will not cause a change in the value stored to \co{x}.
In other words, in this execution there is no semantic dependency
from either \co{y} or \co{z} to \co{x}.
But there \emph{is} a semantic dependency from the \emph{pair}
\{\co{y},~\co{z}\} to \co{x},
because changes to both \co{y} and \co{z}
can cause the value stored in \co{x} to change.
This means that accurate definitions of sdep cannot always rely on
single-variable perturbations;
they must consider changes to multiple variables.
See Appendix~\ref{app:Non-Trivial Semantic Dependencies}
for examples and additional discussion.

Since we no longer regard sdep as always relating a single load to a store,
the notion of a cycle involving sdep appears problematic.
We are forced to change our definition of an OOTA cycle again;
we will say that an execution is an instance of OOTA if in that execution:

\begin{quote}
	There are stores
	$W_i$~($i = 0$, \ldots,~$m$),
	where $W_i$ semantically depends on loads
	$\{R_{i,j}\}$~($j = 0$, \ldots,~$n_i$),
	such that each $R_{i,j}$ reads from one of the $W_k$
	stores in a different thread.
\end{quote}

\noindent
This makes OOTA more complicated than a simple cycle but
we will continue to refer to it as an OOTA cycle out of habit.
Note that this new definition includes and generalizes the earlier
``cycle in sdep $\cup$ rfe'' definition.

\subsubsection{Semantic Dependencies Affected by Cross-Thread Optimizations}
\label{sec:Semantic Dependencies Affected by Cross-Thread Optimizations}

@@

\begin{quote}
\begin{verbatim}
x = y - z;
\end{verbatim}
\end{quote}

\noindent
There appear to be semantic dependencies from \co{y} to \co{x} and from \co{z}
to \co{x}.
However, if the compiler somehow knows that \co{y} is
always equal to \co{z} at this point then there is no semantic dependency;
the compiler can act as if the statement were simply ``\co{x = 0;}''.
We leave aside the question of how the compiler would know this,
given that \co{y} and \co{z} cannot be updated simultaneously
and are subject to change at any time by
other threads (a point we will return to).

@@

Consider the following \co{if} statement:

@@

\noindent
Here there is a semantic dependency from \co{x}, but in some executions
it extends to \co{y} and in others to \co{z}.
This is an example of the load affecting not the value of
a given store, but rather whether or not that store is executed at all.

We have seen several examples showing that semantic dependencies may vary
according to the execution.
This raises the question: What exactly is an execution?
